---
title: "MSDS 6306_402_GroupHW_Wk3"
author: "Anderwald, Frye, Farrar, Vitovsky"
date: "May 25, 2016"
output: 
  html_document:
    keep_md: TRUE
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
```

## Introduction
Our team analyzed the the annual sales data of properties sold in Manhattan, NY.  This data "displays yearly sales information of properties sold in New York City," (http://www1.nyc.gov/site/finance/taxes/property-annualized-sales-update.page).

#Required Packages
The packages **plyr** and **gdata** are required to run commands we performed in organizing the data.  Additionally, we used **knitr** to generate reports from this file, including a BibTex file, and  **repmis** for additional options for reproduceable research.


```{r Required Packages, include=FALSE}
RequiredPackages <- c("plyr", "gdata", "repmis")
lapply(RequiredPackages, require, character=TRUE)
```


#Data File
One can download the file for themselves from the URL https://www1.nyc.gov/assets/finance/downloads/pdf/rolling_sales/neighborhood_sales/manhattan_sales_prices.xls.  We downloaded the most recent data as of the date of our report in May 2016.  After converting the Excel file to a CSV plain text document, we began checking out the data so that we could organize it for analysis. Of course, one can use the gdata package to read the .xls file into R as well. Please change the eval and include commands to "FALSE" or "TRUE" depending on whether you are using Windows or Mac.

```{r ReadCSVonWindows, eval=TRUE, include=TRUE}
ManhattanSales <- read.csv ('~/rollingsales_manhattan.csv', skip=4, header=TRUE)
```

```{r ReadCSVonMac, eval=FALSE, include=FALSE}
ManhattanSales <- read.csv ("rollingsales_brooklyn.xls",pattern="BOROUGH")
```

```{r CheckOutTheData}
head(ManhattanSales)
summary(ManhattanSales)
str(ManhattanSales)
```



